---
title: "ggfacet: facets in ggplot2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`ggfacet` is a simple package providing a wrapper around `ggplot2`'s facetting functions `facet_grid` and `facet_wrap`.

```{r setup}
library(tidyverse)
library(ggfacet)
```

The main interest of `ggfacet` is that it can do this:

```{r, fig.width=5}
facettize(
  
  iris %>%
    mutate(Species = as.numeric(factor(Species))) %>%
    ggplot(aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
    theme_classic() +
    geom_point() +
    theme(legend.position = "none"),
  
  cols = "Species",
  prepend = "alpha==",
  parsed = "Species"
  
)
```

It basically facilitates mathematical notation into facet strips, as the regular `ggplot2` workflow can be a little tedious if you want to parse the strips of the facets as plotmath expressions (e.g. with greek letters).

Other than that, `facettize` is the main function of the package and expects (1) a `ggplot` object to split into facets, (2) a `rows` and/or `cols` argument containing the quoted names of the variables to facet by. For example:

```{r, fig.width=5}
facettize(
  
  iris %>%
    mutate(
      Size = ifelse(Petal.Length > mean(Petal.Length), "large", "small")
    ) %>%
    ggplot(aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
    theme_classic() +
    geom_point() +
    theme(legend.position = "none"),
  
  rows = "Size",
  cols = "Species"
  
)
```
 
 Multiple variables can be combined, in rows or in columns, e.g.
 
```{r, fig.width=5}
facettize(
  
  iris %>%
    mutate(
      Size = ifelse(Petal.Length > mean(Petal.Length), "large", "small")
    ) %>%
    ggplot(aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
    theme_classic() +
    geom_point() +
    theme(legend.position = "none"),
  
  cols = c("Species", "Size")
  
)
```

In order to use `facet_wrap` instead of `facet_grid`, set `wrap = TRUE`. This way the function will not care about whether facetting variables are provided in rows or columns. 

```{r, fig.width=5, fig.height=4}
facettize(
  
  iris %>%
    mutate(
      Size = ifelse(Petal.Length > mean(Petal.Length), "large", "small")
    ) %>%
    ggplot(aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
    theme_classic() +
    geom_point() +
    theme(legend.position = "none"),
  
  rows = "Size",
  cols = "Species",
  wrap = TRUE
  
)
```

Last but not least, you can choose which facetting variables must be parsed as plotmath expressions:

```{r, fig.width=5}
facettize(
  
  iris %>%
    mutate(
      Species = as.numeric(factor(Species)),
      Size = ifelse(Petal.Length > mean(Petal.Length), "large", "small")
    ) %>%
    ggplot(aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
    theme_classic() +
    geom_point() +
    theme(legend.position = "none"),
  
  rows = "Size",
  cols = "Species",
  prepend = "alpha==",
  parsed = "Species"
  
)
```
